---
title: "Regression Models Course Project"
author: "Elizaveta Oginskaya"
date: "June 5, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)
```

## Executive Summary
In this paper a set of car parameters was analysed to find whether MPG depends on transmission type:

* **Is an automatic or manual transmission better for MPG?** Yes, with 95% confidence level we can say that in general MPG is better for automatic transmission.
* **Quantify the MPG difference between automatic and manual transmissions.** Taking into consideration *only* transmission type automatic transmission allows to increase MPG by **7.24** miles per gallon. But there are two other significant parameters: *weight (wt)* and *1/4 mile time (qsec)*. Taking them into account as constant automatic transmission raise MPG only by average **4.30** miles per gallon.

As the report should only include 2 pages of results and annexes can only contain figures, please look at to .rmd file if you have any questions to the code.

## Exploring and Adjusting Data
The datasets contained 32 observations of 11 parameters of different car models (for more information on data please visit http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) *(Annex 1)*
```{r data_adjustment}
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)
```
The original dataset has all variables of numeric class. Before the main part of analysis variables `cyl`, `gear` and `carb` were converted into factors (the code is hidden, but available in *.rmd* file, look it up also for working dataset structure).
```{r exploritory}
str(mtcars)
```

## Statistical Inference
To answer the first question, whether the MPG parameter depends on transmission type or not the t-test is taken, where the null hypothesis assume that the difference in MPG mean for cars with automatic transmission (factor = 1) and cars with manual transmission (factor = 0) is insignificant.
```{r ttest}
t.test(mpg ~ am, data = mtcars)
```
The 95% confidence interval doesn't contain 0 and the p-value is very small, so the null hypothesis should be rejected in favor of the statement that means of two groups of cars are different.

## Regression Model
### Simple Regression Model
To evaluate the difference in MPG for automatic and manual transmissions the simple linear regression model was build. It assumes that MPG depends only on transmission type.
```{r simple_model}
fit1 <- lm(mpg ~ am, data = mtcars)
summary(fit1)$coef
```

```{r simple_rsqr}
summary(fit1)$adj
```

The model shows that the average MPG of cars with manual transmission equals **17.15** and automatic transmission increase MPG by **7.24** to **24.39**. P-values for the both coefficients are relatively small so the both coefficients are significant. Adjusted R-squared value is only 0.338, which means that the model explains only **33.8%** of mpg variation. *(Annex 2)*

### Finding Optimal Regression Model
```{r full_model}
fit10 <- lm(mpg ~ ., data = mtcars)
summary(fit10)$adj
```
For the search of the optimal model three more models were build:
Regression model including all 10 regressors explains more variance of MPG - **77.9%**, but many of its coefficients have relatively large values and appears insignificant (see *.rmd* file)
```{r step}
fits <- step(fit10, direction = 'both', k = log(nrow(mtcars)))
summary(fits)$coef
```

The *stepwise* algorithm by *AIC* was implied to find an optimal fit model. The algorithm found the following best fit model:
```{r step_result}
summary(fits)$call
```
All coefficients of the model except for the intercept turned out to be significant. The adjusted R-squared value for the models equals **0.834**.

On the base of previous model the last model was built by excluding the intercept:
```{r optimal_model}
fitn <- lm(mpg ~ wt + qsec + am - 1, data = mtcars)
summary(fitn)$coef
```
This model has all coefficients significant and explains **98.6%** of MPG variance.
```{r optimal_summary}
summary(fitn)
```

## Nested models
```{r anova}
as.matrix(anova(fit1, fitn))
```
The ANOVA test shows that all three regressors are significant

## Residuals Diagnostics
The residuals plots (Annex ) show that:

* Residuals vs. Fitted plot: points ate randomly distributed, so residuals are IID.
* Normal Q-Q plot: the points mostly follow on the line, so residuals are normally distributed.
* The Scale-Location plot: points distributed in a constant pattern, the variance of residuals is constant.
* Residuals VS Leverage: some distinct points indicates few outliers
```{r residuals}
lvrg <- hatvalues(fitn)
tail(sort(lvrg))

infl <- dfbetas(fitn)
tail(sort(infl[,3]))
```


## Annexes
### Annex 1: Data Pairs
```{r annex1}
par(mfrow = c(1, 1))
pairs(mtcars)
```

### Annex 2: MPG by transmission type:
```{r annex2}
par(mfrow = c(1, 1))
boxplot(mpg ~ am, data = mtcars, xlab = 'Transmission: 0 - manual; 1 - automatic', ylab = 'Miles per Gallon')
```

### Annex 3: Residuals Diagnostics
```{r annex3}
par(mfrow = c(2, 2))
plot(fitn)
```
